---
title: "Page 2"
format: html
editor: visual
---

```{r}
#| echo: false
library(tidyverse)
library(restatapi)
library(knitr)
library(scales)
```

***Why You Should Care***

The Common Agricultural Policy (CAP) is central to the European Union’s efforts to reduce greenhouse gas emissions and combat climate change. Agriculture accounts for a significant share of emissions in Europe, but through CAP funding, farmers are encouraged to adopt sustainable practices such as carbon-efficient crop rotations, precision farming, and the use of renewable energy. These measures not only reduce the environmental footprint of agriculture but also ensure that high-quality food remains available without compromising the planet’s health.

Beyond farming, the CAP supports emission reduction through the restoration of natural ecosystems, such as wetlands and forests, which act as vital carbon sinks. By promoting biodiversity and sustainable land use, the policy contributes to lowering overall emissions while protecting essential resources like soil and water. Every time you choose locally-produced, eco-friendly food, you’re benefiting from CAP-driven initiatives that make agriculture part of the solution to climate change, fostering a greener and more resilient Europe for all.

```{r}
#| include: false
EUghg <- get_eurostat_data("env_air_gge")

summary(EUghg)
```

```{r}
#| echo: false
EUghg %>%
  filter(
    geo == "EU27_2020",  
    airpol == "GHG",     
    src_crf == "CRF3",   
    time %in% c("2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022")  
  ) %>%
  ggplot(aes(x = time, y = values, fill = time)) +  
  geom_bar(stat = "identity", color = "black") +    
  labs(
    title = "EU Greenhouse Gas Emissions
          (2012-2022)",  
    x = "Year",  
    y = "Emissions (Values)",  
    fill = "Year"  
  ) + 
  theme_minimal() +                                  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The chart above displays an aggregate of the Greenhouse Gas emissions related to the agricultural industry in the EU. In the same time period that demand for agricultural products has increased, due to population increase and expansion of trade networks, CAP policies over the past decade have limited and even started to decrease annual emissions related to agriculture across the EU.

```{r}
#| echo: false
EUghg %>%
  filter(
    geo == "EU27_2020",  
    airpol %in% c("CH4", "N2O"),  
    src_crf == "CRF3",  
    time %in% c("2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022")  
  ) %>%
  ggplot(aes(x = time, y = values, fill = airpol)) + 
  geom_bar(stat = "identity", position = "dodge", color = "black") +  
  labs(
    title = "EU Methane (CH4) and Nitrous Oxide (N2O) Emissions (2012-2022)",  
    x = "Year",  
    y = "Emissions (Thousand Tonnes)",  
    fill = "Emission
    Type"  
  ) + 
  scale_y_continuous(labels = comma) +  
  theme_minimal() +                    
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

In this chart we have displayed only Methane and Nitrous Oxide emissions. These are the most prevalent gasses related to agricultural activity and have a disproportionate warming affect on the atmosphere and negative impact on public health. The above data represents a similar mitigating effect on emissions due to CAP initiatives. This shows the global reaching benefits of CAP beyond its immediate objectives.

[Homepage](/Landing.html)

[What is CAP About](/Page1.html)

[CAP and the EU](/Page3.html)

[Who is Affected](/Page4.html)

Sources: “Methane and Climate Change – Global Methane Tracker 2022 – Analysis - IEA.” IEA. https://www.iea.org/reports/global-methane-tracker-2022/methane-and-climate-change#.
