---
title: "Landing Page"
name: "Carter Payne"
format: html
editor: visual
---

## Landing Page

![](images/EU%20FT%20Euro%20Money%20Photo.avif)

Arnold, Martin, Fleming, Sam, & Claire Jones. 2024. "Can Europe’s economy ever hope to rival the US again?." FT Montage/Getty Images. https://www.ft.com/content/93f88255-787b-4c06-849c-f7722c83e8b6.

This website will look at the levels of education and training within the European Union. It will measure different countries' levels of employment to see if there is a correlation between high levels of education and training and high levels of employment within a country.

```{r}
#| include: false
library(tidyverse) 
library(haven) 
library(knitr)
```

```{r}
#| include: false
library(tidyverse) 
library(haven) 
library(knitr)
```

```{r}
#| include: false

library(tidyverse)
library(knitr)
library(restatapi)
library(giscoR)
library(sf)
library(RColorBrewer)
```

```{r}
#| include: false

library(ggplot2)
library(dplyr)
```

```{r}
#| include: false

YoungPeopleinEduandTraining <- get_eurostat_data('edat_lfse_18') 
```

```{r}
#| include: false

dsdYoungPeopleinEduandTraining <-get_eurostat_dsd('edat_lfse_18', verbose=TRUE)
```

```{r}
#| include: false
summary(YoungPeopleinEduandTraining)
```

```{r}
#| echo: false 

library(dplyr)
```

```{r}
#| include: false 

crsLONGLAT <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 

cntrys <- giscoR::gisco_get_countries(
    year = "2020",
    resolution = "3",
    region = c("Europe", "Asia")
) |>
    sf::st_transform(crsLONGLAT)

#names(cntrys)

plot(sf::st_geometry(cntrys))
```

```{r}
#| include: false

YoungPeopleinEduandTraining1 <- YoungPeopleinEduandTraining %>% filter(time == 2022, training=="FE_NFE") 
cntrys_YoungPeopleinEduandTraining1 <- cntrys %>%
  left_join(YoungPeopleinEduandTraining1, by=c("CNTR_ID"="geo"))


head(cntrys_YoungPeopleinEduandTraining1)
```

```{r}
#| echo: false 
crs_lambert <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +datum=WGS84 +units=m +no_defs"

xmin <- -10.66
ymin <- 34.5
xmax <- 45
ymax <- 71

bbox <- sf::st_sfc(
    sf::st_polygon(
        list(
            cbind(
                c(
                    xmin,
                    xmax,
                    xmax,
                    xmin,
                    xmin
                ),
                c(
                    ymin,
                    ymin,
                    ymax,
                    ymax,
                    ymin
                )
            )
        )
    ), crs = 4326
)

lambert_bbox <- sf::st_transform(
    bbox,
    crs = crs_lambert
)

bb <- sf::st_bbox(lambert_bbox)
```

```{r, fig.width=10, fig.height=6}
#| echo: false

cntrys_YoungPeopleinEduandTraining1 %>% 
  ggplot() +
  geom_sf(data = cntrys_YoungPeopleinEduandTraining1,
        mapping = aes(
            fill = values
        ),
        color = "white",
        size = 0.15
    ) +
    coord_sf(
        crs = crs_lambert,
        xlim = c(bb["xmin"], bb["xmax"]),
        ylim = c(bb["ymin"], bb["ymax"]),
    ) + 
  scale_fill_gradientn(
    name = "Percentage in 2022",
    colors = brewer.pal(9, "RdPu"),
    na.value = "grey80"
  ) + labs(title="Percentage of Young People with Formal and Non-Formal Education and Training", caption="Source: Eurostat") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5), 
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        axis.text = element_blank(),
        axis.ticks = element_blank()
        )
```

This [map](https://ec.europa.eu/eurostat/cache/metadata/en/edat1_esms.htm) shows the percentage of young people with formal and non-formal education and training in 2022 in different European countries. The darker the color pink, the higher the percentage of the population that has formal and non-formal education and training. After looking at the map, we can see that Sweden and Finland have the highest percentages of people with formal and non-formal education and training. Eastern European countries, including Bulgaria and Bosnia-Herzegovina, have the lowest percentages of people with formal and non-formal education and training.

Bibliography

Arnold, Martin, Fleming, Sam, & Claire Jones. 2024. "Can Europe’s economy ever hope to rival the US again?." FT Montage/Getty Images. https://www.ft.com/content/93f88255-787b-4c06-849c-f7722c83e8b6.

Eurostat. 2024. Participation Rate of Young People in Education and Training by Sex, Age and Labour Status (incl. NEET rates). Eurostat, the Statistical Office of the European Union. https://ec.europa.eu/eurostat/cache/metadata/en/edat1_esms.htm.

Menu

The following links will redirect you to pages that provide more information on the topic.

[Page 1](/WebsitePage1.html)

[Page 2](/WebsitePage2.html)

[Page 3](/WebsitePage3.html)

[Page 4](/WebsitePage4.html)
